### 2025-03-07 snapshot_return_policy.qmd
thinking about return time for snapshots

---
title: Examining whether the snapshot return mechanism is correct
---

```{python}
import numpy as np
from matplotlib import pyplot as plt
from scipy import stats
from collections import Counter

rng = np.random.default_rng()
```

We want to investigate whether the snapshot return mechanism produces the correct return time distribution.

Currently, because we do not label individuals, all individuals in the same state are treated the same.
This means that individuals in the "at home" bucket are all treated as if they have been there for some time.

We model return (and the overall return rate) in the snapshot model by balancing the number of observed departures (to home) and arrivals (from home).

We can check the concordance of this model with the temporal distribution by writing an individual based-simulation (or adapt our current simulation, where we now also  track marked individuals pretty explicitly)

We will do this by modelling a single link (single we don't care about network effects), and sending individuals through this link indirectly.

```{python}
SNAPSHOT_DURATION = 14
times = np.arange(0, 100)
readmission_duration = lambda size=1: rng.lognormal(np.log(14), 0.5, size=size)

base_population = 1000
base_move_out = 200
```

```{python}
movements_out = np.random.poisson(base_move_out, size=times.shape)

movements_in = np.zeros_like(times)
movement_dist = []
for _t in times: # implictly this is also the index
    move_times = np.round(readmission_duration(movements_out[_t]), 0).astype(int) + _t
    valid_moves = move_times[move_times < np.max(times)]
    movement_dist.extend(valid_moves)
    if len(valid_moves) < 1:
        continue
    move_counter = Counter(valid_moves)
    move_tos = np.array(list(move_counter.keys()))
    move_counts = np.array(list(move_counter.values()))
    movements_in[move_tos] += move_counts
```

Computing the transition matrices
```{python}
at_home = 0
home_numbers = np.zeros_like(times)
for _t in times:
    at_home += movements_out[_t]
    home_numbers[_t] = at_home
    at_home -= movements_in[_t]

inwards_transition_prob = movements_in / home_numbers

outwards_transition_prob = movements_out / base_population
```

```{python}
# create agents
class Agent():
    _place_map = ['hospital', 'home']
    def __init__(self):
        self.place = 0
        self.go_home_time = np.nan
        self.return_hospital_time = np.nan

    def __repr__(self) -> str:
        return f"Agent()"

    def __str__(self):
        return f"Agent [{self.place} {self.go_home_time:.0f} {self.return_hospital_time:.0f}]"

    @property
    def duration(self):
        return self.return_hospital_time - self.go_home_time

    def go_home(self, time):
        self.go_home_time = time
        self.place = 1

    def return_hospital(self, time):
        self.return_hospital_time = time
        self.place = 0

agent_pool = set()
home_pool = set()