### 2025-03-13 link_time_regr.qmd
update some stuff that's been done to analyse the prlbems

---
title: link_time regression
---

```{python}
import igraph as ig
from matplotlib import pyplot as plt
from scipy import stats, optimize
import numpy as np
```

```{python}
B = ig.Graph.Read("concordant_networks/trimmed_static_base_1_threshold_4.graphml")

direct = np.array(B.es['direct_weight'])
indirect = np.array(B.es['indirect_weight'])
link_time = np.array(B.es['link_time'])

plt.loglog(indirect, link_time, '.')
plt.xlabel('Indirect weight')
plt.ylabel('Link Time')
```

```{python}
valid = np.isfinite(link_time)

# transfrom
log_link_time = np.log(link_time[valid])
log_indirect = np.log(indirect[valid])

# estim mean
log_link_mean = np.mean(log_link_time)

residuals = log_link_time - log_link_mean
composed_variable = log_indirect * residuals

est_mean = np.mean(composed_variable)
est_std = np.std(composed_variable)

plt.plot(log_indirect, composed_variable, '.')
```

```{python}
rng = np.random.default_rng()

indirect_values = np.array([250, 30, 40, 70, 110, 100, 40, 5])
composed_draws = rng.normal(est_mean, est_std, size=indirect_values.shape)
drawn_link_times = np.exp(
    (composed_draws / np.log(indirect_values)) + log_link_mean
) / 12 # for time unit consistency

print(drawn_link_times)
```