### 2025-03-13 sim_movement_over_time.qmd

---
title: Movement in Simulations over Time
---

We want to see how much movement there is between home and hospital in each simulation.

```{python}
import h5py
import polars as pl
import numpy as np
from matplotlib import pyplot as plt

rng = np.random.default_rng(2025)
```

```{python}
fh = h5py.File("zero_sims/snapshot/sims_with_30seeds_pseudocap_fixedaxis.h5")
fh_grps = list(fh)

test_index = rng.choice(fh_grps)
test_grp = fh[test_index]

data = {
    k: np.array(v[:], dtype=float)
    for k,v in test_grp.items()
}
```

```{python}
def group_window(arr: np.ndarray, W: int):
    return arr.reshape((-1, W)).sum(axis=1)

WINDOW = 10

plt.plot(data['ts'][:-1:WINDOW], group_window(data['mover_out'].sum(axis=1), WINDOW))
plt.plot(data['ts'][:-1:WINDOW], group_window(data['mover_in'].sum(axis=1), WINDOW))
```

```{python}
fh_static = h5py.File("zero_sims/static/sim_all_30s_fixed_axis.h5")
xi_static = rng.choice(list(fh_static))
data_static = {
    k: np.array(v[:], dtype=float) 
    for k,v in fh_static[xi_static].items()
}

WINDOW = 10

plt.plot(
    data_static['ts'][:-1:WINDOW], 
    group_window(data_static['mover_out'].sum(axis=1), WINDOW)
)

plt.plot(
    data_static['ts'][:-1:WINDOW], 
    group_window(data_static['mover_in'].sum(axis=1), WINDOW)
)
```

... want to look at the departure rate of the first snapshot, and see the distrigubtion. Looks very sparse for snapshot 0. Of 4299 edges, we get 4239 edges with weight < 19, 469 with 0 departure (and thus only direct movement). Note that there are 338^2 possible edges.

```{python}
import igraph as ig

G = ig.Graph.Read("conc_tempo_14_detailed/0000.graphml")
S = ig.Graph.Read("concordant_networks/trimmed_static_base_1_threshold_4.graphml")
```

```{python}
snapshot_df = pl.from_dict(
    {
        'departures': G.es['departures'],
        'arrivals': G.es['arrivals'],
        'source': G.es['source'],
        'target': G.es['target'],
    }
).cast(pl.Int64)

static_df = pl.from_dict(
    {
        'indirect': S.es['indirect_weight'],
        # 'link_time': S.es['link_time'],
        'source': S.es['source_loc'],
        'target': S.es['dest_loc'],
    }
).cast(pl.Int64)

size_df = pl.read_csv("concordant_networks/size_14.csv")

snapshot_df = snapshot_df.join(size_df, left_on='source', right_on='hospital')
static_df = static_df.join(size_df, left_on='source', right_on='hospital')

snapshot_df = snapshot_df.with_columns(dep_rate = pl.col('departures').cast(pl.Float64) / pl.col('estimated_size') / 14)
static_df = static_df.with_columns(dep_rate = pl.col('indirect').cast(pl.Float64) / pl.col('estimated_size') / S['time_span'])

dep_rates = pl.concat([
    snapshot_df
        .with_columns(model=pl.lit('snapshot'))
        .select('dep_rate', 'source', 'target', 'model'), 
    static_df
        .with_columns(model=pl.lit('static'))
        .select('dep_rate', 'source', 'target', 'model')
])

```
```{python}
hist_opts = {
    'histtype': 'step',
    'bins': 31,
    'density': True
}
for name, arr in zip(['snapshot', 'static'], [snapshot_moves, static_moves]):
    plt.hist(arr, **hist_opts, label=name)

plt.yscale('log')
plt.legend()