### 2025-03-27 experiments/small_case/small_case.qmd
Big reshuffle into a more nested directory structure
This is to improve the navgiability and allow for
a more cohesive sharing of code between different parts
of the analysis

---
title: "Small Case: Checking Hitting Times in the Year of Our Lord"
format:
  html:
    code-fold: true
---

# Checking hitting Times: Naive Static Network
```{python Imports} 
import numpy as np
import networkx as nx
import igraph as ig
import polars as pl

import hitting_markov as hmk
import hitting_time_analysis as ht
import hitting_time_multichain as mch
import network_conversion as ntc
import netsim_summariser as summ

from matplotlib import pyplot as plt, colormaps
import seaborn as sns
```

Here we want to check that our simulations and analytical results have parity.

The idea is to setup a small network that has at least one very rare link, and compute the hitting times analytically for N=[1, 5, 10, 30] individuals, and also simulate for those cases.

We generate a small 5-node directed, weighted network.
```{python}
graph_parameters = dict( 
  edges = [
    (0, 1),
    (0, 3),
    (1, 2),
    (1, 3),
    (2, 0),
    (3, 2),
    (3, 4),
    (4, 3),
  ],
  directed = True,
  vertex_attrs = {
    'name': list('ABCDE'),
    'capacity': [100, 120, 200, 100, 30],
  },
  edge_attrs = {
    'weight': [
      300, # (0, 1),
      50, # (0, 3),
      50, # (1, 2),
      100, # (1, 3),
      150, # (2, 0),
      120, # (3, 2),
      60, # (3, 4),
      10, # (4, 3),
    ]
  }
)
G = ig.Graph(**graph_parameters)

G.write_graphml("small_case/graph.graphml")

fig, ax = plt.subplots()
ig.plot(
  G,
  target = ax,
  vertex_label=G.vs['name'],
  vertex_color='lightcyan',
  vertex_size=50,
  edge_label=G.es['weight'],
  edge_background='white',
  edge_label_dist=1.7,
)
```

We compute Q, the transition matrix.
We extract T, the base transition matrix, by allowing defaults of using the vertex indices as the ordering, since we have the sizes attached to the graph.
```{python}
TIMESPAN = 28
T = ntc.transition_matrix_from_graph(
  graph=G,
  scaling_per_node=G.vs['capacity'],
  global_scaling=TIMESPAN,
  adjacency_attribute='weight'
)
Q = hmk._Q_mat_naive(T)
```

This is our ground truth, the hitting time for N=1, as extracted from the linear system that results.
```{python}
def draw_ecdf(values, ax=None, **kwargs):
  value_arr = np.sort(values.flatten())
  ecdf_quantiles = np.linspace(0, 1, len(value_arr), endpoint=False)

  if ax is None:
    fig, ax = plt.subplots()
  ax.step(value_arr, ecdf_quantiles, **kwargs)
  return ax
```

```{python}
hitting_time_one,_ = ht.compute_static_hitting_times(Q)
draw_ecdf(hitting_time_one)
```

Here we use a scaling=100 to get proper convergence of the integrals. This value is relatively arbitrary, but it must be >1 since there are slow species that ruin our time.
```{python}
hitting_time_numerical = mch.compute_all_multichain_hittings(1, Q=Q, scaling=100)
ax = draw_ecdf(hitting_time_numerical)
```

These are our numerical base results. We "hope" they are correct (insofar as numerical error)
```{python}
chain_range = [1, 5, 10, 30]
ht_numers = {
  n_chains: mch.compute_all_multichain_hittings(n_chains, Q=Q, scaling=100)
  for n_chains in chain_range
}

fig, ax = plt.subplots()
for k, v in ht_numers.items():
  draw_ecdf(v, ax=ax, label=f"N={k}")
plt.legend()
```

We can then simulate on this naive static network.
We will offload this work to another script, mostly because the simulations take a significant amount of time (and we only want to run it once)

[script](small_case_sims.py)

Similarly, we offload the construction of the metrics.

```{python}
sim_archive = "small_case/simulations.h5"
summariser = summ.Summariser(sim_archive)
metrics = summariser.metrics(ncpus=6, no_move=True, drop=[])
export_file = "small_case/metrics.parquet"
metrics.write_parquet(export_file)
```

Here, we construct the empirical CDF
```{python}
# metrics = pl.read_parquet("small_case/metrics.parquet")
hitting_time_cols = pl.selectors.starts_with('hitting_time_')
aggregated_metrics = metrics.unpivot(
  on=hitting_time_cols,
  index=['nchains','seed',],
  variable_name='target',
  value_name='hitting_time',
)
max_value_plus_one = aggregated_metrics.select(pl.col('hitting_time').max()).item() + 1
cleaned_metrics = aggregated_metrics.fill_nan(max_value_plus_one).fill_null(max_value_plus_one)

fig, axz = plt.subplots(nrows=2)
axs = axz.flatten()

for ax in axs:
  sns.ecdfplot(
    cleaned_metrics, 
    x='hitting_time', 
    hue='nchains', 
    palette=colormaps.get('jet'),
    ax=ax
  )

  for k, v in ht_numers.items():
    draw_ecdf(v, ax=ax, label=f"N={k}")

axs[1].set_xlim(-10, 100)

```

We see relatively good agreement here.

### Aside: the Bug

We detected a bug in this work, where the silent graceful failure of numpy.random.Generator.multinomial causes a bug where the transition probabilities that were not normalised correctly earlier cause high probability moves to the "last" node.

This causes the ECDFs to clump, and be differentiated by their hitting probabilities.

```{python}
bug_metrics = pl.read_parquet("small_case/metrics_prebug.parquet")
hitting_time_cols = pl.selectors.starts_with('hitting_time_')
aggregated_bugmetrics = bug_metrics.unpivot(
  on=hitting_time_cols,
  index=['nchains','seed',],
  variable_name='target',
  value_name='hitting_time',
)
cleaned_bugmetrics = aggregated_bugmetrics.fill_nan(max_value_plus_one).fill_null(max_value_plus_one)

ax = sns.ecdfplot(cleaned_bugmetrics, x='hitting_time', hue='nchains', palette=colormaps.get('jet'))
```

# Checking Hitting Times: Improved Static Network

Here, we will want to extend the construction above where we split the graph edge weights between two projections.

```{python}
bigraph_edge_attrs = {
   'direct_weight': [
      50, # (0, 1),
      20, # (0, 3),
      10, # (1, 2),
      30, # (1, 3),
      40, # (2, 0),
      20, # (3, 2),
      20, # (3, 4),
      5, # (4, 3),
    ],
    'indirect_weight': [
      250, # (0, 1),
      30, # (0, 3),
      40, # (1, 2),
      70, # (1, 3),
      110, # (2, 0),
      100, # (3, 2),
      40, # (3, 4),
      5, # (4, 3),
    ],
    # see link_time_regr.qmd
    'link_time': [
      7.52, # (0, 1),
      8.82, # (0, 3),
      20.35, # (1, 2),
      7.10, # (1, 3),
      11.51, # (2, 0),
      10.49, # (3, 2),
      10.99, # (3, 4),
      5.46, # (4, 3),
    ]
}

bigraph_parameters = dict(graph_parameters)
bigraph_parameters['edge_attrs'] = bigraph_edge_attrs

B = ig.Graph(**bigraph_parameters)
B.write_graphml("small_case/bigraph.graphml")

fig, ax = plt.subplots()
ig.plot(
  B,
  target = ax,
  vertex_label=G.vs['name'],
  vertex_color='lightcyan',
  vertex_size=50,
  edge_label=list(zip(B.es['direct_weight'], B.es['indirect_weight'], B.es['link_time'])),
  edge_background='white',
  edge_label_dist=1.7,
)
```

So we have split the graph into two layers, a direct alyer, and an indirect layer. I haven't checked the direct/indirect split in the real graph, but we will assume for now that this is an appropriate split.
We have derived link times by empirical resampling of the real graph link times, dividing by 12 to get smaller link times appropriate for a shorter global time scale, and then assigning the sampled link times by matching order of graph edge weight -> link time (higher weight gets shorter time)

We can compute the transition matrix Q for this by performing an expansion of the home states.
```{python}
B_size_map = {nd['name']: nd['capacity'] for nd in G.vs}
Qb = ht.extract_delayed_static_Qmat(
  B, 
  scaling=TIMESPAN,
  size_mapping=B_size_map,
  ordering_key='name', 
  value_key='capacity',
)
```

We can, again, compute the $N=1$ chain case.

```{python}
hitting_time_imprvst_numer = mch.compute_subset_multichain_hittings(
  Nchains = 1,
  Q = Qb.todense(),
  index_subset = slice(Q.shape[0]),
  scaling = 100,
)
ax = draw_ecdf(hitting_time_numerical, label='naive static')
ax = draw_ecdf(hitting_time_imprvst_numer, ax=ax, label='improved static')
ax.legend()
```

And the $N \in \{1, 5, 10, 30\}$ cases
```{python}
ht_imprv_numers = {
  n_chains: mch.compute_subset_multichain_hittings(
    n_chains, 
    Q=Qb.todense(), 
    index_subset=slice(Q.shape[0]), 
    scaling=100
  )
  for n_chains in chain_range
}

fig, ax = plt.subplots()
for i, (k, v) in enumerate(ht_imprv_numers.items()):
  if k == 1: continue
  draw_ecdf(v, ax=ax, label=f"N={k}", color=f"C{i}")
  draw_ecdf(ht_numers[k], ax=ax, label=f"N={k} [naive]", linestyle='dashed', color=f"C{i}")
plt.legend()
```

Let us also look at the scripts that run the simulations.

```{python}
home_sim_archive = "small_case/home_sims.h5"
home_summariser = summ.Summariser(home_sim_archive)
home_metrics = home_summariser.metrics(ncpus=6, no_move=True, drop=[])
home_export_file = "small_case/home_metrics.parquet"
home_metrics.write_parquet(home_export_file)
```

Plotting ...

```{python}
aggregated_home_metrics = home_metrics.unpivot(
  on=hitting_time_cols,
  index=['nchains','seed',],
  variable_name='target',
  value_name='hitting_time',
)
max_home_value_plus_one = aggregated_home_metrics.select(pl.col('hitting_time').max()).item() + 1
cleaned_home_metrics = aggregated_home_metrics.fill_nan(max_home_value_plus_one).fill_null(max_home_value_plus_one)

fig, axz = plt.subplots(nrows=2)
axs = axz.flatten()

for ax in axs:
  sns.ecdfplot(
    cleaned_home_metrics, 
    x='hitting_time', 
    hue='nchains', 
    palette=colormaps.get('jet'),
    ax=ax
  )

  for k, v in ht_imprv_numers.items():
    draw_ecdf(v, ax=ax, label=f"N={k}")

axs[1].set_xlim(-10, 100)
```

Comparing observed hitting time distributions between the naive and improved static network models

```{python}
fig, axs = plt.subplots(
  nrows=2, ncols=2,
  sharex=True, sharey=True,
  layout='constrained')
axz = axs.flatten()

for ax, nchain in zip(axz, chain_range):
  sns.ecdfplot(cleaned_metrics.filter(pl.col('nchains') == nchain), x='hitting_time', ax=ax, label='Naive')
  sns.ecdfplot(cleaned_home_metrics.filter(pl.col('nchains') == nchain), x='hitting_time', ax=ax, label='Improved')
  ax.set_title(f"Nchains={nchain}")
ax.legend()
```