### 2025-03-27 notebooks/compare_hitting_times.qmd

---
title: Comparing hitting times filtering by seed/target
---

The main aim is to construct a function that defines a quantifiable metric of whether a hitting time 
eCDF is below or above another.

This is to determine whether or not the hitting time distribution for X model is always "above" Y model; and also the extended question of whether X > Y > Z always. (or almost always)

```{python}
import numpy as np
import polars as pl
from scipy import stats
from matplotlib import pyplot as plt
import seaborn as sns
import itertools

from typing import Mapping
```

```{python}
# going to steal from aggregated_compare.ipynb

metrics_archives = {
    'temporal': './zero_sims/temporal/metrics_30s.parquet',
    'static': './zero_sims/static/metrics_30s_fa.parquet',
    'snapshot': './zero_sims/snapshot/metrics_30s_fa.parquet',
    'naive_static': './zero_sims/naive_static/metrics_30s.parquet',
}

raw_metrics = {
  label: pl.read_parquet(archive) for label, archive in metrics_archives.items()
}

hitting_time_prefix = "hitting_time_"
hitting_time_columns = pl.selectors.starts_with(hitting_time_prefix)
```

```{python}
def add_ecdf(df: pl.DataFrame, name='eCDF'):
  augmented_df = (
    df
    .sort('hitting_time', nulls_last=True)
    .with_row_index(name=name)
    .with_columns(pl.col(name) / pl.col(name).max())
  )
  return augmented_df

aggregated_metrics = dict()
for model, metrics in raw_metrics.items():
  agg_df = (
    metrics
    .unpivot(
      on=hitting_time_columns,
      index='seed',
      variable_name='target_name',
      value_name='hitting_time',
    )
    .with_columns(
      target=(
        pl.col('target_name')
        .str.strip_prefix('hitting_time_')
        .str.to_integer()
      )
    )
    .drop('target_name')
    .sort('hitting_time', nulls_last=True)
    .with_row_index(name='CDF')
    .with_columns(
      pl.col('CDF') / pl.col('CDF').max()
    )
  )
  aggregated_metrics[model] = agg_df
```

```{python}
sns.scatterplot(aggregated_metrics['temporal'], x='hitting_time', y='CDF', marker='.')
```

```{python}
def hitting_time_curve(df, x='hitting_time', y='CDF', ax=None, **kwargs):
  if ax is None:
    _, ax = plt.subplots()

  hitting_time = df[x].to_numpy()
  ecdf = df[y].to_numpy()

  ax.plot(hitting_time, ecdf, **kwargs)

  return ax
```

```{python}
hitting_time_curve(aggregated_metrics['temporal'])
```

```{python}
def partial_ecdf_plot(dfs: Mapping[str, pl.DataFrame], filter_fun: pl.Expr):
  ax = None
  for key, df in aggregated_metrics.items():
    aug_df = add_ecdf(df.filter(filter_fun), name="partial eCDF")
    ax = hitting_time_curve(aug_df, y='partial eCDF', ax=ax, label=key)
  ax.legend()
  return ax
```

```{python}
partial_ecdf_plot(aggregated_metrics, pl.col('seed').eq(0))
```

I think the idea here will be to compare N-quantiles of the distribution curve

```{python}
base_quantiles = pl.from_dict(
  {'quantile': np.linspace(0, 2900, 30)}
)

def quantile_metric(df: pl.DataFrame, quantiles=base_quantiles, cdf_col='CDF'):
  searched = quantiles.join_asof(
    df, 
    left_on='quantile',
    right_on='hitting_time', 
  ).select('quantile', cdf_col)

  return searched

def get_partial_quantiles(df: pl.DataFrame):
  # compute partial eCDF
  cdf_col = 'eCDF'
  aug_df = add_ecdf(df, name=cdf_col)
  # search quantiles
  return quantile_metric(aug_df, cdf_col=cdf_col)
```

## Single seed test

Testing for seed = 42
```{python}
test_seed = 42

qmets = {model: get_partial_quantiles(df.filter(pl.col('seed').eq(test_seed))) for model, df in aggregated_metrics.items()}

qmet_agg = pl.concat([
  df.with_columns(model=pl.lit(model))
  for model, df in qmets.items()
])

qmet_matrix = qmet_agg.pivot(on='model', index='quantile', values='eCDF')
```

```{python}
sns.scatterplot(qmet_agg, x='quantile', y='eCDF', hue='model')
```

First attempt: Cramer-von Mises . The problem here is that these goodness of fit criteria are two-sided, and we want a one-sided criterion.
```{python}
comb_symbol = "__x__"
qmet_cramer = qmet_matrix.with_columns(
  *(
    ((pl.col(col1) - pl.col(col2))).alias(f"{col1}{comb_symbol}{col2}")
    for col1, col2 in itertools.combinations(aggregated_metrics.keys(), 2)
  )
).select(pl.selectors.contains(comb_symbol)).sum().to_dicts()[0]
```

Second attempt was to compute P( X > Y ), but we have censored distributions, so this is hopeless.

```{python}
qmet_simple = qmet_matrix.with_columns(
  *(
    (pl.col('naive_static') - pl.col(col)).alias(f'naive_minus_{col}')
    for col in aggregated_metrics.keys()
  )
).select(
  pl.selectors.starts_with('naive_minus_').name.map(
    lambda s: s.removeprefix('naive_minus_')
  )
).sum().to_dicts()[0]

qmet_order = sorted(qmet_simple, key=qmet_simple.get)
```

### Whole of seed

```{python}
def filter_seed(df, seed):
  return df.filter(pl.col('seed').eq(seed))

def filter_target(df, target):
  return df.filter(pl.col('target').eq(target))

def q_simple_order(seed=None, target=None):

  if seed is None and target is None:
    raise ValueError("Must specify at least one of seed or target")
  elif seed is None:
    filterf = lambda df: filter_target(df, target)
  elif target is None:
    filterf = lambda df: filter_seed(df, seed)

  qmets = {
    model: get_partial_quantiles(filterf(df)) 
    for model, df in aggregated_metrics.items()
  }

  qmet_agg = pl.concat([
    df.with_columns(model=pl.lit(model))
    for model, df in qmets.items()
  ])

  qmet_matrix = qmet_agg.pivot(on='model', index='quantile', values='eCDF')

  qmet_simple = qmet_matrix.with_columns(
    *(
      (pl.col('naive_static') - pl.col(col)).alias(f'naive_minus_{col}')
      for col in aggregated_metrics.keys()
    )
  ).select(
    pl.selectors.starts_with('naive_minus_').name.map(
      lambda s: s.removeprefix('naive_minus_')
    )
  ).sum().to_dicts()[0]

  qmet_order = tuple(sorted(qmet_simple, key=qmet_simple.get))

  return qmet_order
```

```{python}
all_q_seed = {i: q_simple_order(seed=i) for i in aggregated_metrics['naive_static'].select(pl.col('seed').unique()).to_series().to_list()}

all_q_targ = {i: q_simple_order(target=i) for i in aggregated_metrics['naive_static'].select(pl.col('target').unique()).to_series().to_list()}
```

```{python}
levels = list(itertools.permutations(aggregated_metrics.keys()))
levels_map = {s: i for i,s in enumerate(levels)}

level_q_seed = [levels_map[all_q_seed[k]] for k in sorted(all_q_seed.keys())]
# not using targ keys, since it includes the shitty 149
level_q_targ = [levels_map[all_q_targ[k]] for k in sorted(all_q_seed.keys())] 
```

```{python}
for lnm, llq in zip(('seed', 'target'), [level_q_seed, level_q_targ]):
  plt.figure()
  plt.title(lnm)
  plt.plot(llq, '.')
  yt = plt.gca().get_yticks()
  plt.gca().set_yticklabels(
    [levels[int(t)] if t < len(levels) else '' for t in yt]
  );
```

Seeing if there is correlation between the ordering and the (estimated) size of the hospital. We know that this estimated size has an artificial lower bound of 5.
```{python}
sizes = pl.read_csv("concordant_networks/size_14.csv").sort('hospital').with_row_index(name='index')

level_df = pl.from_dict({
  'index': sorted(all_q_seed.keys()),
  'seed_order': level_q_seed,
  'targ_order': level_q_targ,
  'seed_str': [all_q_seed[k] for k in sorted(all_q_seed.keys())],
  'targ_str': [all_q_targ[k] for k in sorted(all_q_seed.keys())],
})

level_df = level_df.join(sizes, on='index')
```

```{python}
sns.scatterplot(
  level_df.select('index', 'seed_order', 'estimated_size'),
  x='estimated_size',
  y='seed_order',
)
```

```{python}
jj_level_df = level_df.select('seed_order', 'estimated_size').group_by(pl.all()).len().sort('len', descending=True).with_columns(pl.col('len').log(10).alias('loglen'))
```

```{python}
nonstandard_order_df = level_df.filter(pl.col('seed_order').eq(21))
xhosp = nonstandard_order_df.filter(pl.col('estimated_size').eq(pl.col('estimated_size').max())).select('index').item()

partial_ecdf_plot(aggregated_metrics, pl.col('seed').eq(xhosp))
```

```{python}
nonstandard_hosps = nonstandard_order_df.select('index').to_series().to_list()
for hosp in nonstandard_hosps:

  naive_hosp = add_ecdf(aggregated_metrics['snapshot'].filter(pl.col('seed').eq(hosp)), name='partialCDF')

  maxv = naive_hosp.drop_nulls(pl.col('hitting_time')).select(pl.col('partialCDF').max()).item()

  print(hosp, maxv)
```