### 2025-03-27 notebooks/explore_tau_over_time.qmd

---
title: Exploring the distribution of $\tau$ over time
execute:
  echo: false 
---

We want to explore the distribution of $\tau$, the interadmission time, as a function of the time that an individual leaves hospital.

```{python}
import polars as pl
from matplotlib import pyplot as plt
import seaborn as sns
from pathlib import Path

import graph_importer as gim
import tempograph_extractor as tge

basepath = Path(".")

sns.set_context('talk')
sns.set_style("ticks")
sns.set_palette("colorblind")
```

We load in the 1-day discretisation for temporal fidelity

```{python}
#| echo: false
CACHE = True
if not CACHE:
  G = gim.make_graph((basepath / "concordant_networks/temponet_1_365.lgl").absolute().as_posix())
  edge_df = tge.extract_edge_df(G)
  tau_series = tge.pull_series_tau_dist(edge_df)
  tau_dists = tge.compute_tau_dists(tau_series)
  clean_tau_dists = tau_dists.drop_nans()
else:
  clean_tau_dists = pl.read_csv(basepath / "concordant_networks/analyses/tau_dists.csv").drop_nans()
```

```{python}
fig, axs = plt.subplots(nrows=2, sharex=True)
axz = axs.flatten()
sns.scatterplot(clean_tau_dists, x='time', y='slope', ax=axz[0])
sns.scatterplot(clean_tau_dists, x='time', y='intercept', ax=axz[1])
```