### 2025-03-27 reports/sketch.qmd
Update report

---
title: "DRAFT: Towards a paper on patient movement in the Victorian hospital system"
subtitle: "an exercise in trying to avoid reality"
author: "X-- xXx THE AUTHORS xXx --X"
filters:
  - tikz 
format: html
---

- Introduction
  - CPE is a concern in Victoria, and we want to manage it and prevent it from spreading
    - A good way to understand potential patterns of spread is to understand how patients move between facilities in the system
  - Inter-facility patient movement models for studying CPE
    - mostly static networks, or projected to static networks
    - but temporal fluctuations can potentially affect the dynamics
  - Temporal network methodology is relatively well established
    - but little on systems where information/contact leaves the system for periods of time
  - Our aim is to see whether or not this temporal fidelity is needed for this context
    - Can we get similar movement dynamics with simpler approximations
- Setting and Data Sources
  - Standard spiel
- Models and Methods
  - "Standard" model of

:::{#fig-mineandonlymine}
```{.tikz}
%%| format: svg
%%| filename: transfer_diagram
%%| img-width: 900
\usetikzlibrary{arrows.meta, positioning, shapes.geometric}

\begin{tikzpicture}[
    node distance=2cm, 
    hospital/.style={draw, circle},
    >=latex
  ]

  \begin{scope}[every node/.style = {hospital}]
  \node (A1) {A};
  \node (B1) [below=of A1] {B};
  \node (A2) [right=of A1] {A};
  \node (A3) [right=of A2] {A};
  \node (B2) [below=of A2] {B};
  \node (B3) [below=of A3] {B};
  \end{scope}

  \draw[->, dashed] (A1) edge[bend left] node [above] {indirect self-transfer} (A3);
  \draw[->, dashed] (A1) edge (B2);
  \draw[->, dashed] (A1) edge node [right] {indirect transfer} (B3);
  \draw[->] (A1) edge  node[above] {continuity} (A2);
  \draw[->] (A2) edge  (A3);
  \draw[->] (A1) edge node[left] {direct transfer}  (B1);
\end{tikzpicture}
```

Patient movements from hospital A at a given time: (a) a direct transfer to another facility B at the same time, (b) an indirect transfer to another facility B at a different time, going home in between, (c) an indirect self-transfer, where the patient is discharged home, but readmits at the same facility at a later time, (d) a continuity step where the patient stay at a given facility
:::

Here, we are interested in two processes:

1. the process that discharges patients from hospital
2. the process that readmits patients to hospital

We can characterise each by a random variable that represents the time they spend in the preceding state.
These random variables -- $s$ for the discharge process and $\tau$ for the readmission process -- are modelled differently in each of the different network representations.

| Network Representation | Model of $s$ (per edge) | Model of $\tau$ (per edge) |
|:------:|:------------------:|:------------------:|
| Naïve Static | Exponential | 0 |
| Static with Return Home | Exponential | Exponential |
| Static Snapshots | Exponential, varies over time | Difference of Exponentials, varies over time |
| Temporal | Exponential, varies with time | Empirical |

{{< placeholder >}}
... need an image here for the empirical distribution of $s$.

1. Results
  - Temporal network model yields slower and less diverse spread than other network models.
  - A naïve static network model vastly overestimates the spread 
  - Some improvement can be made by using networks and models that incorporate more temporal fidelity, but they do not capture the magnitude of change in the temporal network model

```{python}
#| echo: false
import netsim_summariser as summ
import polars as pl
import seaborn as sns
import numpy as np
from matplotlib import pyplot as plt
from scipy import stats

sns.set()
sns.set_palette('colorblind')
```

```{python}
metrics_archives = {
    'temporal': './zero_sims/temporal/metrics_30s.parquet',
    'static': './zero_sims/static/metrics_30s_fa.parquet',
    'snapshot': './zero_sims/snapshot/metrics_30s_fa.parquet',
    'naive_static': './zero_sims/naive_static/metrics_30s.parquet',
}

processed_results = {}

for label, archive in metrics_archives.items():
        processed_results[label] = pl.read_parquet(archive)

hitting_time_columns = pl.selectors.starts_with('hitting_time_')

agg_metrics = dict()
hitting_time_dists = dict()
for model, metrics in processed_results.items():
    agg_met = (metrics
                   .unpivot(
                     on=hitting_time_columns, 
                     index='seed', 
                     variable_name='target', 
                     value_name='hitting_time'
                   )
                   .with_columns(target_seed=(
                       pl.col('target')
                       .str.strip_prefix('hitting_time_')
                       .str.to_integer()
                   ))
                   .drop('target')
    )
    agg_metrics[model] = agg_met
    hitting_time_dists[model] = np.sort(agg_met.select('hitting_time').to_series().to_numpy())

analytical_expected_hitting_naive = pl.read_csv("./hitting_time_analysis/numer_hitting_time_30indv_chains.csv")
ana_E_hitT_naive = np.sort(analytical_expected_hitting_naive.select('hitting_time').to_series().to_numpy().flatten())

plt.figure()
for model, dist in hitting_time_dists.items():
    plt.plot(dist, np.linspace(0, 1, len(dist), endpoint=False), label=model)

xlim = plt.gca().get_xlim()
ylim = plt.gca().get_ylim()

plt.plot(ana_E_hitT_naive, np.linspace(0, 1, len(ana_E_hitT_naive)), 
         label='naive\nanalytical', linestyle='dashed')

plt.xlim(xlim)
plt.ylim(None, 0.9)

# plt.ylim(1e-3, None)
plt.xlabel('hitting time')
plt.ylabel('ECDF')

plt.legend(loc='lower right')
```