### 2025-04-01 reports/model.qmd
Update model

---
title: "Section: Model of Stochastic Processes and Network Models"
filters:
  - tikz 
format: html
execute:
  echo: false
---

Notations

- $u,v$ for locations; discrete, usually subscript
- $s,t$ for times; continuous, usually function inputs/variables
- $T, \omega, \Omega$ for characteristic times
- $N$ as a size variable
- $x, y, z$ are state variables; often functions of location and time

:::{.hidden}
$$
\newcommand{\dd}[2][t]{\frac{\mathrm{d}#2}{\mathrm{d}#1}}
\newcommand{\ceil}[1]{\lceil#1\rceil}
$$
:::

## General Sketch

:::{#fig-general-diagram}
```{.tikz}
%%| format: svg
%%| filename: movement_general_case
%%| img-width: 800
%%| fig-cap: general movement processes
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, shapes.arrows}

\begin{tikzpicture}[
  node distance=2cm and 4cm,
  >=latex
]
  \tikzset{
    squarx/.style= {
      regular polygon,
      regular polygon sides=4
    }
  }
  \tikzset{
    house/.style= {
      single arrow,
      shape border rotate=90,
      single arrow head extend=0pt,
      single arrow tip angle=120,
      minimum height=0.75cm,
      minimum width=0.75cm
    }
  }

  \node[draw, squarx] (A) {$u$};
  \node (X) [right=of A] {};
  \node[draw, squarx] (B) [right=of X] {$v$};
  \node[draw, house] (H) [below=of X] {};
  \node[draw, circle] (D) [above=of X] {$\emptyset$};

  \draw[->] (A) edge node[above]{$d_{uv}(s, t)$}  (B);
  \draw[->] (A) edge node[above left]{$\zeta_u(s)$} (D);
  \draw[->] (A) edge node[below left]{$\eta_{uv}(s,t)$} (H);
  \draw[->] (H) edge node[below right]{$\rho_{uv}(s,t)$} (B);
\end{tikzpicture}
```

Sketch of processes modelled
:::

Here, $u$ and $v$ are locations (potentially identical locations), $s$ and $t$ are times. We take that $t \geq s$.
$\emptyset$ represents patients that are not observed to return to the hospital system after their discharge from location $u$. We assume the associated hazard for this process ($\zeta_u(t)$) to be known for all locations $u$ at all times $t$: the modelling of this process is the same for all network models, and will be neglected in the following.

Further, we define the function $w_{uv}(s, t)$ to be the number of individuals that are observed to discahrge from $u$ at time $s$ and readmit at time $t$. We allow a maximum of one of the pairs $u,v$ and $s,t$ to be reflexive, i.e. $u=v$ or $s=t$, but not both.

We will also assume that we will have quantities $N_u$ that represent the "size" of a location $u$. We assume that all locations $u$ will have $N_u$ individuals in them at all times, and that the observed mvoements $w_{uv}(s, t)$ can be thought of as being scaled by $N_u$.

The governing equations can be roughly expressed as a function of the hazards of each process.

$$\begin{aligned}
\lambda(u(s) \to \emptyset) &= \zeta_u(s)\\
\lambda(u(s) \to v(t)) &= d_{uv}(s, t)\\
\lambda(u(s) \to z_{uv}(t)) &= \eta{uv}(s, t)\\
\lambda(z_{uv}(t) \to v(t)) &= \rho_{uv}(s, t)
\end{aligned}$$

Note that

$$ \lambda(U \to V) = \lambda $$

represents that in some time step $\Delta t$, the number of individuals that move from state $U$ to state $V$ is distributed as

$$ \mathrm{Poisson}(\lambda x_U \Delta t) $$.

### Naïve Static Model

For both the naïve and improved static network models, we have that the observations of $w$ are taken over a time period of $T_\Sigma$.

In the naïve static model , we have that:

$$d_{uv}(s,t) = d_{uv} = \frac{\sum_{s,t} w_{uv}(s,t)}{T_\Sigma N_u}$$
$$\eta_{uv}(s,t) = \eta_{uv} = 0$$
$$\rho_{uv}(s,t) = \rho_{uv} = 0$$

That is, all movements are instantaneous, and the rate of movement is the mean rate of movement over the entire observation period.

### Improved Static Model

For the improved static model, we choose some threshold value $\omega$ that delineates direct and indirect transfers.

We will have:

$$d_{uv}(s,t) = d_{uv} =  \frac{\sum_{s,t: (t-s) < \omega} w_{uv}(s,t)}{T_\Sigma N_u}$$
$$\eta_{uv}(s,t) = \eta_{uv} = \frac{\sum_{s,t: (t-s) \geq \omega} w_{uv}(s,t)}{T_\Sigma N_u}$$
$$\rho_{uv}(s,t) = \rho_{uv} = \left[ \frac{\sum_{s,t: (t-s) \geq \omega} (t-s) w_{uv}(s,t)}{\sum_{s,t: (t-s) \geq \omega} w_{uv}(s,t)}  \right]^{-1}$$

Here, we model the rate of indirect transfers $\eta$ similarly to the rate of direct transfers  $d$, by counting the average rate of observed movements. For the rate of readmission, $\rho$, we use the inverse of the mean readmission duration/delay.

### Snapshot Model

For the snapshot model, we choose a snapshot duration $\omega$. This defines the threshold duration of an indirect transfer.

This introduces an new intermediary state variable $z'$ that contains the individuals that would enter $z$ within the duration of a given snapshot $[t, t+\omega]$, so that they do not immediately readmit to their next hospital.
This alters the indirect transfer processes to

$$\begin{aligned}
\lambda(u(s) \to {z'}_{uv}(t)) &= \eta_{uv}(s, t)\\
\lambda({z'}_{uv}(t) \to z_{uv}(t)) &= \delta(t \mathrm{ mod } \omega)\\
\end{aligned}$$

where $\delta(\cdot)$ is the Dirac delta function.

Alongside this, we also compute from the data, the number of patients "at home" at a given time, in order to inform the denominator for the process from $z'$ to $z$. We denote this quantity $H_{uv}(t)$

We will have that:

$$d_{uv}(s,t) = d_{uv}(s) = \frac{\sum_{t: (t-s) < \omega} w_{uv}(s,t)}{\omega N_u}$$
$$\eta_{uv}(s,t) = \eta_{uv}(s) = \frac{\sum_{t: (t-s) \geq \omega} {w_{uv}(s, t)}}{\omega N_u}$$
$$\rho_{uv}(s,t) = \rho_{uv}(t) = \frac{\sum_{s: (t-s) \geq \omega} {w_{uv}(s, t)}}{\omega H_{uv}}$$

### Temporal Model

For the temporal network, we make a choice of the time discretisation $\omega$. That is, events that occur within a time window $[\tau, \tau+\omega)$ are collapsed and aggregated to occur at time $\tau$. Over the observation (and simulation) period, we will have a series of time window boundaries $\{\tau_0, \tau_1, \dots\}$. We denote the smallest value in that set larger than a given time $t$ to be $\ceil{t}_\omega$

For the readmission process, we assume that individuals that would readmit within a given time window $[\tau, \tau+\omega)$ do so uniformly.

This yields:

$$d_{uv}(s, t) = \frac{\sum_{t: (t-s) < \omega} w_{uv}(s, t)}{\omega N_u}$$
$$\eta_{uv}(s, t) = \frac{\sum_{t: (t-s) \geq \omega} w_{uv}(s, t)}{\omega N_u}$$
$$\rho_{uv}(s,t) = \frac{1}{\ceil{t}_\omega - t}$$

The readmission process thus has a uniform hazard.