### 2025-04-09 notebooks/rare_path.qmd
Update rare path for example graph diagrmans and analysis

---
title: Rare Pathing
---

## Direct rare pathing
```{python}
import igraph as ig
import numpy as np
from echidna import network_simulation as netsim

from matplotlib import pyplot as plt
import seaborn as sns
import polars as pl
```

```{python}

G = ig.Graph(
    n=4, 
    edges=[
        (0, 1),
        (1, 2),
        (2, 0),
        # (0, 3),
    ], 
    directed=True, 
    edge_attrs={'weight':[
        1000,
        1000,
        1000,
        # 0,
    ]}
)

R = ig.Graph(
    n=4, 
    edges=[
        # (0, 1),
        # (1, 2),
        # (2, 0),
        (0, 3),
    ], 
    directed=True, 
    edge_attrs={'weight':[
        # 0,
        # 0,
        # 0,
        100,
    ]}
)
```

```{python}
#| label: fig-graphs
fig, axs = plt.subplots(ncols=2)
axz = axs.flatten()

ig.plot(
    G,
    target=axz[0],
    edge_label=G.es['weight'],
    layout='circle',
)

ig.plot(
    R,
    target=axz[1],
    edge_label=R.es['weight'],
    layout='circle',
)

axz[0].set_xlabel('regular')
axz[0].spines[:].set_visible(True)
axz[1].set_xlabel('rare')
axz[1].spines[:].set_visible(True)
```

```{python}
N_size = np.array([3000, 3000, 3000, 100]).reshape((-1, 1))

T_G = np.array(G.get_adjacency('weight').data) / N_size

T_R = np.array(R.get_adjacency('weight').data) / N_size

T_stack = np.stack([T_G[:,:,np.newaxis], T_R[:,:,np.newaxis], *[T_G[:,:,np.newaxis] for _ in range(998)]], axis=2)

T_ave = np.squeeze(np.mean(T_stack, axis=2))
```

```{python}
## STATIC
step_size = 0.2
gamma_static = T_ave.sum(axis=1, keepdims=True) # (-1,1)
W_ave = T_ave / gamma_static
W_ave[np.isnan(W_ave)] = 0.0

static_sim = netsim.Simulation(
    full_size = N_size,
    parameters = {
        'beta': 0.0,
        'gamma': gamma_static,
        'prob_final_stay': np.zeros_like(gamma_static),
        'transition_matrix': W_ave,
    },
    dt = step_size,
)

static_sim.fixedseed(0, 1000)
```

```{python}
## SNAPSHOT
gamma_normal = T_G.sum(axis=1, keepdims=True)
gamma_rare = T_R.sum(axis=1, keepdims=True)

W_normal = T_G / gamma_normal
W_rare = T_R / gamma_rare

W_normal[np.isnan(W_normal)] = 0.0
W_rare[np.isnan(W_rare)] = 0.0

snapshot_sim = netsim.SnapshotNaive(
    N_size, 
    parameters={
        'beta': 0.0,
        'transition_matrix': W_normal,
        'gamma': gamma_normal,
        'prob_final_stay': np.zeros_like(gamma_normal),
        'transition_matrices': [W_normal, W_rare, W_normal],
        'gammas': [gamma_normal, gamma_rare, gamma_normal],
    },
    timings=[0, 5, 8],
    dt=step_size,
)

snapshot_sim.fixedseed(0, 1000)
```

run and analyse the snapshot stuff
```{python}
N_SIMS = 1000

snapshot_histories = []
for it_ in range(N_SIMS):
    snapshot_sim.reset(soft=True)
    snapshot_sim.simulate(30, nostop=True)
    snapshot_histories.append(snapshot_sim.history)
```

```{python}
timez = [snapshot_sim.ts[x] 
         for hist in snapshot_histories 
         for x in np.flatnonzero(hist[-1,:])[:1]
        ]

plt.hist(timez)
```

run and analyse static
```{python}
static_histories = []
for it_ in range(N_SIMS):
    static_sim.reset(soft=True)
    static_sim.simulate(30, nostop=True)
    static_histories.append(static_sim.history)
```

```{python}
static_timez = [static_sim.ts[x] 
         for hist in static_histories 
         for x in np.flatnonzero(hist[-1,:])[:1]
        ]
```
```{python}
#| label: fig-cumhist
plt.xlabel('hitting time')
# plt.ylabel('count')
# plt.hist(static_timez, histtype='step', bins=31, label='static')
# plt.hist(timez, histtype='step', bins=4, label='snapshot')
plt.ylabel('cumulative count')
plt.hist(static_timez, cumulative=True, histtype='step', bins=31, label='static')
plt.hist(timez, cumulative=True, histtype='step', bins=4, label='snapshot')
plt.legend()
# plt.yscale('log')
```

## Indirect rare pathing

```{python}
S = ig.Graph(
    n=4, 
    edges=[
        (0, 1),
        (1, 2),
        (2, 0),
        # (0, 3),
    ], 
    directed=True, 
    edge_attrs={
        'weight': [
            1000,
            1000,
            1000,
            # 0,
        ]
    }
)
```