### 2025-04-14 reports/report.qmd
split old (pre aggd hiting time) and new

---
title: "Temporality is Important for Networks of Hospital Patient Transfers"
authors:
    - name: David Wu
    - name: Michael Lydeamore
    - name: Andrew Stewardson
    - name: Tjibbe Donker
    - name: Nic Geard
    - name: "******"
date: last-modified
format:
    html: default
    pdf:
        include-in-header:
            - text: |
                \usepackage{mathtools}
        papersize: a4
bibliography: refs.bib
csl: plos.csl
filters:
    - tikz 
jupyter: python3
execute:
    echo: false
    warning: false
    error: false
    eval: false
callout-appearance: simple
---

```{python}
import numpy as np
import polars as pl
from matplotlib import pyplot as plt, colors
import seaborn as sns
import h5py
import itertools
import re

from pathlib import Path

# setup printing formats for tables
pl.Config.set_tbl_hide_column_data_types(True)
pl.Config.set_tbl_hide_dataframe_shape(True)

sns.set_theme(style='ticks')
sns.set_palette('colorblind')
sns.set_context('paper')
```

# Key Points

- There is a significant amount of patient movement in the Victorian hospital system where the patient returns home for a substantive amount of time between discharge from one hospital and subsequent admission at another hospital.
- This "indirect" movement has implications on the rate of spread of a hypothetical contagion through the hospital system. It causes dispersal of patients to be slower than expected.
- We can recover some of this behaviour by introducing an additional layer in the network to represent indirect movement.

<!-- # Glossary

| symbol | name | description |
|--------|------|-------------|
| v | vertex | |
| e | edge | |
| i | index of vertex | |
| j | index of other vertex | |
| t | time | |
| x | location | hospital |
| W | bin width | discretisation of the temporal network |
| $\gamma$ | discharge | rate of discharge |
| N | num | size of something |
| k | hitting time | (expected) |
| $\lambda$ | rate | of an event / process |
| T | time period | of the _entire_ simulation |

define a vertex $v_i = (x_i, t_i)$ made up of a location and time resp. -->

# Introduction

Antimicrobial resistance poses a great threat to human health and development. [@murray_global_2022]
Globally, it poses a large burden, operationally and economically, on hospital systems.
In Australia, we see [significant] economic burden from AMR.
this impact could become more significant if we had incursion and subsequent establishment of new [strains] of AMR. 
One X of interest is carbapenemase-producing enterobacteriacea (CPE). 
Australia has an endemic strain of CPE -- IMP4 -- but this is at a relatively low level. 
This already causes some level of burden, and introduced strains would further multiply this.

Understanding of the spread of AMR tharity of the temporality]]]]].

# Models and Methods

## Setting and data sources

Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of $N_x = 338$ hospital campuses, which vary in size, case mix and services.

Hospital admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
In particular, admissions data from 1 January 2011 to 9 December 2020 (3631 days) was used in this study. All Victorian hospital admissions during this study period are included. The total number of admission records is 26876787.
This data was aggregated into time-discretised temporal networks (described below), and exported from the database; further analysis as presented in this paper is performed on external machines with no access to identifiable protected health information.

Some exploratory analysis of this data was done in by [@lydeamore_carbapenemase-producing_2024].

## Existing Network Modelling {#sec-network-modelling}

There has been some existing work that analyse hospital systems using the framework of network analysis. 
They define a network consisting of hospitals as nodes, connected by edges that have weights representative of the amount of patient movement between those hospitals.
Using these networks, hospitals and links that are important to mrough hospital systems has been studied in the past as a network phenomenon [@donker_patient_2010]. 
The movement of patients through the hospital systems induced by referrals can be represented as edges between hospital nodes. 
However, these networks are usually presented as static, where movement over a large period of time has been aggregated to form a single network. 
It is known, especially in other applications, that temporal networks can display different behaviour than the static equivalents. 
In this paper, we investigate these differences using the Victorian hospital system as a baseline for comparison. 
In particular, we investigate the impacts of the inclusion of modelling the period of time that patients spend at home between admissions, and the effect of approximating the granulovement around the hospital system by analysing measures of centrality and clustering.
<!-- Thus the network can encode a representation of the movement between hospitals over a given time period. -->

A weighted, directed network can be used to construct a Markov chain that in turn can be analysed or simulated to understand the dynamics of a spreading infectious disease.

For each edge $e_{ij}$ with a number of observed movements $w_{ij}$ over some time period $\tau$, 
we can construct a reaction that has a hazard function $\lambda_{ij} = w_{ij} / \tau$. This implicitly constructs a process that has inter-event times (i.e. times between distinct patient movements) that are exponentially distributed $\text{Exp}(\lambda_{ij})$.

This type of construction is

:::{.callout-note} 
[Stuff about previous implementations of this idea] 
:::

## Network Models for Temporal Fidelity

Static networks provide a very interpretable model of influence of nodes in the system: network measures allow for some semblance of understanding of aggregate aspects of the network and its components. However, dynamic models derived from those static networks miss important dynamics and temporal qualities of the system they are modelling.
It is well-known that temporal networks exhibit different behaviour to analogous static networks: they have lower connectivity ....

In the context of the spread of AMR between healthcare facilities, these differences can lead to misleading results about the dynamics of the [disease], and related control mechanisms and policies.
Using the detailed linkage data from the VAED, we can constrct a relatively detailed temporal network, which can then be projected in various network representations with different temporal fidelity.
We roughly consider three aspects of the dynamics:

1. the representation of the time patients spend outside the healthcare system (i.e. at home)
1. the time-varying number of transfers (direct and indirect) between facililties
1. the distribution of the inter-admission times for patients that are discharged and subsequently readmit to the system

{{< include _model.qmd >}}

These models have distinct advantages and disadvantages: the "simpler" models are more amenable to analysis, but the more "complex" models represent more fidelity.
For example, the snapshot network model allows us to analyse each static snapshot as its own graph, and then subsequently aggregate those results as a time series; this is a much smaller algorithmic complexity compared to any analogous analysis that could be done on the full temporal network.

In the extreme case, we can analyse a toy model in order to demonstrate the difference between the naïve and improved static network models.

Consider a two-node network, where patients transfer from hospital A to hospital B. Suppose over some time period $T$ there are $n_d$ direct transfers and $n_i$ indirect transfers.
In the naïve model, the expected hitting time is equal to $\frac{T}{n_d+ n_i}$.
In the improved model, we suppose we also have the mean inter-admission time of $\tau$.
Then the expected hitting times $k$ can be solved by a linear system [@Norris_Markov_Chains_1997]:

$$
\begin{bmatrix}
\frac{T}{n_d+n_i} & -\frac{T}{n_d} & -\frac{T}{n_i} \\
0 & 1 & 0 \\
0 & -\frac{1}{\tau} & \frac{1}{\tau}
\end{bmatrix} \begin{bmatrix} k_A \\ k_B \\ k_H \end{bmatrix}
= \begin{bmatrix} 1 \\ 0 \\ 1\end{bmatrix}
$$

which solves to give $k_A = \frac{T}{n_d+n_i}\left(1 + \frac{n_i}{T} \tau \right)$.
Somewhat expectedly, the expectation of the time taken to reach hospital B from hospital A is longer by a factor related to the number of indirect movements and their average inter-admission period.

# Simulation Experiments

Beyond simple models, there is vanishingly small possibility of finding tractable analystical expressions for the hitting times.

## Experimental Setup

We initialise the state of a simulation by seeding 30 individuals in a single facility.
We then simulate the system forwards in time until a fixed end time.

Each starting facility is looped over to be the starting facility, and realisations are acquired 20 times for each starting facility.
This gives pairs of hitting times between the (known) starting facility and target facility (which we get for all non-starting facilities by the end of the simulation realisation)

<hr />