### 2025-06-09 reports/diagrams/network_diags/temporal.tex
meaningless small updates

\documentclass[tikz, margin=2mm]{standalone}

\usetikzlibrary{
    arrows.meta, 
    positioning, 
    shapes.geometric, 
    shapes.arrows,
    decorations.pathmorphing,
}

\begin{document}

\begin{tikzpicture}[scale = 0.20]

\tikzset{
    wavy/.style={
        decorate,
        decoration={
            snake,
            pre length=2mm,
            post length=2mm
        }
    }
}

\node at (19.0,-15.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](A0){$u_s$};
\node at (19.0,-47.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](B0){$v_s$};
\node at (61.0,-15.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](A1){$u_t$};
\node at (61.0,-47.0) [draw,solid,minimum width=1.6 cm,minimum height=1.6 cm,inner sep=0](B1){$v_t$};

% direct transfers
\draw [-stealth,solid](A0.247)|-++(0,-1)-|(B0.113);
\draw [-stealth,solid](B0.67)|-++(0,1)-|(A0.293);
\draw [-stealth,solid](A1.247)|-++(0,-1)-|(B1.113);
\draw [-stealth,solid](B1.67)|-++(0,1)-|(A1.293);

% time consistency
\draw [-stealth,wavy](A0.east)--(A1.west);
\draw [-stealth,wavy](B0.east)--(B1.west);

% indirect
\draw [out=45,in=135,-stealth,dashed,bend left=30](A0.45) to (A1.135);
\draw [out=315,in=225,-stealth,dashed, bend right=30](B0.315) to (B1.225);
\draw [out=315,in=135,-stealth,dashed](A0.315) to (B1.135);
\draw [out=45, in=225,-stealth,dashed](B0.45) to (A1.225);

\end{tikzpicture}

\end{document}