### 2025-06-23 reports/diagrams/new-networks/static.tex

\documentclass[tikz, margin=2mm]{standalone}

\usetikzlibrary{
    arrows.meta, 
    positioning, 
    shapes.geometric, 
    shapes.arrows,
    decorations.pathmorphing,
}

\begin{document}

\begin{tikzpicture}

    \node at (0.0, 0.0) (NORTHWEST){};
    \node at (10.0, 10.0) (SOUTHEAST){};

    \node at (3.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A){$u$};
    \node at (3.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B){$v$};
    \node at (7.0, 5.0) [draw,circle,minimum width=1 cm,inner sep=0](H){$h_{uv}$};

    \draw [out=225,in=135,-stealth,solid](A.225)to(B.135);
    %\draw [out=45,in=315,-stealth,solid](B.45)to(A.315);
    \draw [out=0,in=270,-stealth,densely dashed](B.0)to(H.270);
    \draw [out=90,in=0,-stealth,densely dashed](H.90)to(A.0);

\end{tikzpicture}

\end{document}