### 2025-06-23 reports/diagrams/new-networks/temporal.tex

\documentclass[tikz, margin=2mm]{standalone}

\usetikzlibrary{
    arrows.meta, 
    positioning, 
    shapes.geometric, 
    shapes.arrows,
    decorations.pathmorphing,
}

\begin{document}

\begin{tikzpicture}

    \node at (0.0, 0.0) (NORTHWEST){};
    \node at (16.0, 10.0) (SOUTHEAST){};

    \node at (4.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A){$u$};
    \node at (4.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B){$v$};

    \node at (12.0, 8.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](A2){$u$};
    \node at (12.0, 2.0) [draw,solid,minimum width=1 cm,minimum height=1 cm,inner sep=0](B2){$v$};

    \draw [out=225,in=135,-stealth,solid](A.225)to(B.135);
    \draw [out=0,in=180,-stealth,densely dashed](B.0)to(A2.180);

    \draw [loosely dashed,very thick](8.0, 0.0)--(8.0, 10.0);

\end{tikzpicture}

\end{document}